---
title: Identifying hot spots
author: Justin, Natalia, Alex
date: Jul 31, 2020
output:
  html_document:
    code_folding: show
    toc: true
    toc_float: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=14, fig.height=8, echo=TRUE, eval=TRUE, cache=TRUE,
                      warning=FALSE, message=FALSE)
library(ggplot2)
library(covidcast)
library(tidyverse)
library(dplyr)
load_all("/home/shyun/repos/covidcast/R-packages/covidcast") ## Load the covidcast pacakge
```

## Simple first try


Download data.

```{r}
## Use case:
data_sources = c("fb-survey", "fb-survey", "indicator-combination")
                ## "hospital-admissions",
                ## "ght")
signals = c("smoothed_hh_cmnty_cli", "smoothed_cli", "confirmed_7dav_incidence_num")
           ## "smoothed_adj_covid19",
           ## "smoothed_search")
signals = data.frame(data_sources = data_sources, signals = signals)
suppressMessages({
  mat = covidcast_signals(signals,
                          start_day = "2020-06-01", end_day = "2020-07-31")
})
mat = mat %>%
  as_tibble() %>%
  dplyr::select(geo = geo_value, time =time_value, val = value, signal = signal)

## ## Pick a few geo values for illustration
## mat = mat_orig %>% subset(geo %in% c("01010", "01020"))
## geos = (mat_orig$geo %>% unique() )[10:12]

## ## Make the three geo regions' plots
## mat %>% subset(geo %in% geos) %>%
##   group_by(geo) %>%
##   ggplot(aes(x = time, y = val, col=signal))  + geom_line(size=1) + geom_point(size=2) +
##   facet_wrap(~geo, scales = 'free')
```

Now pick geo values by the highest rises in case counts from June to July.

```{r}
## Helper to calculate rise from June to July
howmuchrise <- function(val, time){
 mean(val[time >= "2020-7-26"]) - mean(val[time <= "2020-6-06"])
}

## Sort all the rises
rises = mat %>%
  group_by(geo) %>% 
  subset(signal == "confirmed_7dav_incidence_num") %>% 
  dplyr::summarise(rise = howmuchrise(val, time)) %>%
  arrange(desc(rise))

## See the numerical summary
rises %>% dplyr::select(rise) %>% unlist() %>% summary()

## Histogram
ggplot(rises, aes(x=rise)) + geom_histogram() %>%
  ggtitle("Distribution of mean(last week of July) - mean(first week of June)") +
  theme(plot.title = element_text(size = 20, face = "bold"))

## Sorted values as a line plot
ggplot(cbind(ind = 1:nrow(rises), rises), aes(x=ind, y=rise)) + geom_line() +
  geom_vline(xintercept=20, col='red', size=1) +
  ggtitle("Sorted mean(last week of July) - mean(first week of June)") +
  theme(plot.title = element_text(size = 20, face = "bold"))

## Pick the 20 highest rising counties
topn = 40
geos = rises %>%  slice_head(n = topn) %>% dplyr::select(geo) %>% unlist()

## library(genlasso)
loess_smooth <- function(val){
  dat = data.frame(ind=1:nrow(val), val)

  predict(loess(val~ind, dat, span=0.3), newx = 1:nrow(val))

  ## ## cubic trend filter
  ## a = genlasso::trendfilter(val, ord=1,maxsteps=)
  ## cv = genlasso::cv.trendfilter(a)
  ## suppressMessages({predict(a, lambda=cv$lambda.1se)$fit})
}

library(maps)
data(county.fips)
conv = county.fips
conv = conv %>% rename(geo=fips)
conv = conv %>% as_tibble %>% mutate(geo=sprintf("%05d", geo))
mat = mat %>% full_join(conv)

totalcensus::convert_fips_to_names(
  FIPs,
  states = NULL,
  geo_header = "STATE",
  in_states = NULL
)

## Make the three geo regions' plots
mat %>% subset(geo %in% geos) %>%
  group_by(geo, signal) %>% 
  ## group_by(signal) %>% 
  mutate(val=scale(val)) %>%
  ## mutate(val2=loess_smooth(val)) %>%
  ## mutate(val2=c(NA,diff(val2))*4)%>%
  ungroup() %>%
  subset(signal != "smoothed_cli") %>%
  group_by(polyname) %>%
  ggplot(aes(x = time, y = val, col=signal))  +
  geom_line(aes(x=time, y=val), size=2) +
  geom_point(aes(x=time, y=val), size=1) +
  geom_hline(yintercept=0, col='grey', size=1) +
  facet_wrap(~polyname, scales = 'free') 
```



**Ryan said:** For getting examples of the two questions that I wrote above (basically,
positive examples, and negative examples), let me suggest we proceed in this
way, a bit more automatically:

- Identify the top say $20$ or so counties that had the biggest percent increase
(absolute, or relative?) in case incidence rate, when you compare the first week
of June to the last week of July. **Done**

- For each, just look at the plots, and see how many count as positive examples,
and negative examples. **Done**

- I guess you can just consider the FB community signal for now (forget the FB
individual one). **Done**

- When you make these plots, there’s some ambiguity/trickiness here in the
scaling: depending on how you scale each signal, you could be interpreting
things visually differently.  I’d think concretely about how strongly this is
affecting things. **Addressing this next**

