---
title: Predicting hotspots one week ahead
author: Justin, Natalia 
date: Jul 31, 2020
output:
  html_document:
    code_folding: show
    toc: true
    toc_float: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=14, fig.height=8, echo=TRUE, eval=TRUE, cache=TRUE,
                      warning=FALSE, message=FALSE)
library(ggplot2)
library(covidcast)
library(tidyverse)
```

## Simple first try

First, download some data (two signals, and case counts).

```{r}
load_all("/home/shyun/repos/covidcast/R-packages/covidcast") ## Sorry for the
                                                             ## absolute path,
                                                             ## you need to load
                                                             ## the package
                                                             ## directly

## Use case:
data_sources = c("doctor-visits",
                "fb-survey",
                "indicator-combination")
                ## "hospital-admissions",
                ## "ght")
signals = c("smoothed_adj_cli",
           "smoothed_cli",
           "confirmed_7dav_incidence_num")
           ## "smoothed_adj_covid19",
           ## "smoothed_search")
signals = data.frame(data_sources = data_sources, signals = signals)
suppressMessages({
mat = covidcast_signals(signals,
                        start_day = "2020-04-01", end_day = "2020-05-07")
})
```

Then, make this into wide format so that every row is a unique time and geo-value.

```{r}
mat = mat %>%  select(geo_value, time_value, data_source, value)
mat_wide = mat %>% pivot_wider(names_from = c("data_source"),values_from="value")
mat_wide %>% print()
```

Now, we form a response vector whose values are 0 or 1 based on whether a 25% increase occurred in 1 day.

```{r}
## Helper function for adding a new response column to every submatrix by geo-value.
onelag = function(mt, numlag){
  numlag = 1
  case = mt$"indicator-combination"
  mt$lagdif = c(rep(NA, numlag),
                sapply((1+numlag):length(case), function(ii){
                  inds1 = ii
                  inds2 = ii-1
                  ((mean(case[inds1]) - mean(case[inds2])) / mean(case[inds2]) ) > 1.25
                }))
  mt
}

## Now, transform a set of wide
mat_new = mat_wide %>% ##subset(geo_value %in% c("01000", "01001")) %>% 
  group_by(geo_value) %>%
  group_modify(onelag)

## Isolate attention to a few regions
geos = mat_new %>% select(geo_value) %>% unique()  %>% unlist()
mat_small = mat_new %>% subset(geo_value %in% geos[2:10])

## Plot the surrounding points? Clunky code hereon..
inds = which(mat_small$lagdif == 1)
inds = sapply(inds, function(ind)ind + (-20):20) %>% as.numeric()
inds = inds[inds>0]

myplot <- function(mat, dummy){
  plot(y=mat$"indicator-combination", x=mat$time_value, type='l' )
  which(mat$lagdif) %>% print()
  abline(v=which(mat$lagdif))
}

mat_small = mat_small %>% rename(case="indicator-combination")

mat_small %>%
  group_by(geo_value) %>%
  ggplot(aes(x = time_value, y = case)) +
  geom_line(colour = "black") + facet_wrap(~geo_value)
```

## Next up:
- Every row of the data is a unique geo/time; make the time variable correspond to the time of the response.
- Better response variable indicating a hotspot; 25 percent increase against a 3 day average.
- Other idea: response is whether 25% increase in cases occurs (0/1) in k=3 weeks? <-- vary k?
- Covariates are various other features. <-- add more things.
- Actually fit a logistic model.
